---
title: "Practical Machine Learning"
author: "Mauricio Santelices"
date: "15-02-2021"
output: html_document
---

## Summary

We will build a model to predict if a user of a user of a device such as Jawbone Up, Nike FuelBand, and Fitbit is lifting weights.

The dataset consists on 5 classes:

 - Class A: Subject lifting weights exactly according to the specification
 - Class B: Throwing the elbow to the front
 - Class C: Lifting the dumbbell halfway
 - Class D: Lowering the dumbbell halfway
 - Class E: Throwing the hips to the front

For more details, the URL of the original dataset is: <http://groupware.les.inf.puc-rio.br/har>


## Getting data

```{r}
if(!file.exists("pml-training.csv")) {
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv", method = 'curl')
}
dataset <- read.csv("pml-training.csv", na.strings = c("NA", ""))

if(!file.exists("pml-testing.csv")) {
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv", method = 'curl')
}
validation <- read.csv("pml-testing.csv")
```


## Data preprocessing

Import packages and set seed:

```{r}
library(caret)
library(randomForest)

set.seed(1756)
```


## Data partition

70% of the data will be the training set and the rest will be the testing set.

```{r}
inTrain <- createDataPartition(y = dataset$classe, p = 0.7, list = FALSE)
training <- dataset[inTrain, ]
testing <- dataset[-inTrain, ]
```

Eliminate the NA entries.

```{r}
naCols <- sapply(training, function(x) {sum(is.na(x))})
colsNA <- names(naCols[naCols > 0])
training <- training[, !names(training) %in% colsNA]
names(training)

training <- training[, !names(training) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

Remove NAs in the validation set:

```{r}
naCols <- sapply(validation, function(x) {sum(is.na(x))})
colsNA <- names(naCols[naCols > 0])
validation <- validation[, !names(validation) %in% colsNA] 

validation <- validation[, !names(validation) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```


And for the testing set:

```{r}
naCols <- sapply(testing, function(x) {sum(is.na(x))})
colsNA <- names(naCols[naCols > 0])
testing <- testing[, !names(testing) %in% colsNA]

testing <- testing[, !names(testing) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```


## Building the model

We will build a Random Forest:

```{r acc}
training$classe <- as.factor(training$classe)
testing$classe <- as.factor(testing$classe)

model <- randomForest(classe ~ .,   data=training, ntree = 50)
predictions <- predict(model, testing)
confusionMatrix(predictions, testing$classe)
modelAcc <- confusionMatrix(predictions, testing$classe)$overall[[1]]
```

Our model is `r modelAcc` accurate.

Now, we will predict the unknown classes of the validation set.

```{r}
predictions <- predict(model, validation)
predictions
```